Draft Exploration

### 2020-04-21 draft_exploration.Rmd
mapped draftees by state

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
library(DataCombine)
library(RColorBrewer)
```

@@

mlb_career <- mlb_start %>%
  left_join(mlb_end, by = "mlb_ID") %>%
  mutate(mlb_length = mlb_end - mlb_start + 1) %>%
  select(mlb_ID, mlb_end, mlb_length)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
library(DataCombine)
library(RColorBrewer)
library(usmap)
```

@@

mlb_career <- mlb_start %>%
  left_join(mlb_end, by = "mlb_ID") %>%
  mutate(mlb_length = mlb_end - mlb_start + 1) %>%
  select(mlb_ID, mlb_start, mlb_end, mlb_length)

@@

# I filtered the master data for signed high school picks before 2012 to make it
# likely they are either in the majors or not in an organization. I grouped by
# draft region, which for high school picks is the abbreviation of their home
# state. I then summarized the total drafted, total to make it to MLB, and the
# percent to MLB. I then joined the summarized data to the statepop data from
# usmap package to get the state names for mapping. I filtered out states with
# fewer than ten draft picks and plotted the data.

statepop <- statepop

state_mlb <- master %>%
  filter(source == "H" & draft_numb == 1 & signed == "Y") %>%
  filter(year <= 2012) %>%
  mutate(mlb_made_numeric = ifelse(high_level == "MLB", 1, 0)) %>%
  group_by(draft_region_id) %>%
  summarize(percent_mlb = sum(mlb_made_numeric) / n(),
            drafted = n(),
            mlb = sum(mlb_made_numeric))

state_mlb <- state_mlb %>%
  left_join(statepop, by = c("draft_region_id" = "abbr"))

state_mlb <- state_mlb %>%
  filter(!is.na(full)) %>%
  mutate(full = reorder(full, drafted),
         big_state = ifelse(draft_region_id %in% c("CA", "TX", "FL"), 1, 0))

plot_usmap(data = state_mlb %>%
             filter(drafted >= 50),
           values = "percent_mlb", color = "white") + 
  scale_fill_viridis_c() +
  theme(legend.position = "right")

```{r}
state_mlb %>%
  filter(big_state == 0) %>%
  ggplot(aes(full, drafted, color = percent_mlb)) +
  coord_flip() +
  geom_point() +
  scale_y_continuous(breaks = c(0, 50, 100, 200, 300)) +
  scale_color_viridis_c() +
  labs(
    x = "State",
    y = "Number of Players Drafted",
    title = "Number of High School Players Drafted by State",
    subtitle = "From 1965-2012 - Outliers: (California: 2249, Texas: 627, Florida: 1069)",
    color = "MLB Percent"
  )
```
